---
title: "Problem Set 1 Q1"
author: "Yicheng Ding"
date: "9/30/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
# Part a

```{r}
install.packages("opendatatoronto")
install.packages("tidyverse")
install.packages("visdat")
install.packages("skimr")

library(opendatatoronto)
library(tidyverse)
library(visdat)
library(skimr)

COVID_19_Cases_in_Toronto <- list_package_resources("https://open.toronto.ca/dataset/covid-19-cases-in-toronto/")
COVID_19_Cases_in_Toronto_Statistics <- COVID_19_Cases_in_Toronto %>%
  get_resource()
```

This data set is about the COVID-19 Cases in Toronto. As we all know, COVID-19 is the most serious epidemic at this time. By analyzing it, we can clearly know its trend. It is interesting to know if there is a relationship between the confirm cases and age, gender. Also, we can know what is the most popular source of infection. This data set contain huge amount of data, so it might be easy for us to find the pattern.

# Part b

```{r}
COVID_19_Cases_in_Toronto_Statistics %>%
  ggplot(mapping = aes(x=Outcome)) +
  geom_histogram(stat = "count")
```

From the graph, we can find that most of the COVID-19 Cases in Toronto have been resolved. There are no more than 2500 active cases exist now. Comparing the fatal cases and the resolved cases, we can find that the fatality rate is not too high.

# Part c

```{r}
COVID_19_Cases_in_Toronto_Statistics %>%
  select(`Age Group`) %>%
  pull() %>%
  table() %>%
  pie()
```

From the graph, we can find that most positive cases are concentrated in the age of group 20-29 years, 30-39 years, 40-49years, 50-59 years and 60-69 years. The positive cases of 19 and younger are not too much which means young people have stronger resistance to the COVID-19.

# Part d

```{r}
COVID_19_Cases_in_Toronto_Statistics %>%
  select(`Client Gender`) %>%
  pull() %>%
  table() %>%
  barplot()

COVID_19_Cases_in_Toronto_Statistics %>%
  select(`Source of Infection`) %>%
  pull() %>%
  table() %>%
  pie()
```


  Base on the Client Gender bar plot, we can find that overall it seems there is no relationship between the confirm cases and the gender. However, the proportion of female confirm cases is slightly larger than male confirm cases, this might because the immunity of women is not as strong as that of men.
  Base on the source of infection pie plot, we can find that close contact and community are the most common source of infection overall. This shows the importance of keeping social distance, which the Toronto government did pretty well recently.
  Although the situation of COVID-19 has got better recently, the government should still be take action to prevent a large-scale outbreak

# Part e

I do not have any references.